\section{Blue Variables}\label{blueVariables}

Blue treats special text as special Blue variables. Below is a list of
those variables.

\begin{longtable}[]{@{}ll@{}}
\toprule
Variable & Value\tabularnewline
\midrule
\endhead
\textless{}TOTAL\_DUR\textgreater{} & Duration of the generated score
from the timeline (Available in Global SCO text area.)\tabularnewline
\textless{}INSTR\_ID\textgreater{} & Replaces with instrumentId; if the
instrument is a named instrument, value is quoted. Generally used when
creating notes.(Available within instruments and some
SoundObjects.)\tabularnewline
\textless{}INSTR\_NAME\textgreater{} & Replaces with instrumentId; if
the instrument is a named instrument, the value is not quoted. Generally
used when working with ORC code to give a variable a unique ID, i.e.
"gk\_var\textless{}INSTR\_NAME\textgreater{}".(Available within
instruments and some SoundObjects.)\tabularnewline
\textless{}RENDER\_START\textgreater{} & The start time of rendering.
Does not take into account time warping of score. (Available in Global
SCO text area.)\tabularnewline
\textless{}RENDER\_START\_ABSOLUTE\textgreater{} & The start time of
rendering. Takes into account time warping of score. (Available in
Global SCO text area.)\tabularnewline
\textless{}PROCESSING\_START\textgreater{} & The time value when
always-on effects instruments need to start. Calculated as the time
value of the first soundObject that is not a FrozenSoundObject or
Comment. (Available in Global SCO text area.)\tabularnewline
\bottomrule
\end{longtable}
